### 2017-02-08 rebuttal.tex
JSB submitted version

\documentclass{article}

\usepackage[utf8]{inputenc}
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%
%opening
\title{Response to Reviewers' Comments}

\begin{document}

\maketitle

We thank both reviewers for their valuable feedback which was extremely
helpful. We have incorporated their suggestions into the paper to improve both content and
readability of our paper. Below we first comment on some overall improvements and changes that 
we made to the algorithm, followed by point-by-point reply to the reviewers' comments.

\section*{Improvements to the Manuscript}
\subsection*{Details about using conjugate gradient to solve eqn. 9 }
While $L$ in eqn. 9 is PSD, the new effective operator in the LHS of eqn. 11 is not necessarily PSD in general.
In order to use conjugate gradient, we solve the system $S^{-1}L(S^{-1}\Sigma_S S^{-1})S^{-1} = S^{-1}MS^{-1}-I$, where
$\Sigma_S = S \Sigma S$, in which the operator acting on $\Sigma_S$ in the LHS is PSD. $\Sigma$ is then obtained
from the estimated $\Sigma_S$. We note that this change did not effect whatsoever the output of the algorithm,
but we wanted the reviewers to be aware of this small change.

\subsection*{Outlier detection}
The reviewers raised a very important question of how outliers are affected by the denoising procedure.
This is a very interesting question, and we were curious to see
the results ourselves. We find that outlier images such as pure noise images have
relatively low contrast after denoising using CWF. So a simple linear classifier
based on the image contrast can be employed for outlier detection. More complicated situations
such as outliers from a different class of a heterogeneous dataset are difficult to distinguish
with denoising. We have added an experiment in section 3.7 for detecting pure noise images as outliers, to demonstrate
how CWF can be extended to build a classifier for outlier detection.

\section*{Response to Reviewer 1}
\begin{itemize}
 \item Page 2: ``The first is a heuristic approach known as phase flipping". I 
would remove the word heuristic, since correcting for the sign, although not 
perfect, is already a well-justified correction. \\
\textbf{\textit{ As per the reviewer's suggestion, we have dropped the word `heuristic'.}}

 \item In the theoretical justification of the method, one of the main 
assumptions is that the $X_i$ and $\xi_i$ coefficients are iid random variables. 
$X_i$ are not independent of each other (see C.O.S. Sorzano, J. Vargas, J. Otón, 
V. Abrishami, J.M. de la Rosa-Trevín, S. del Riego, A. Fernández-Alderete, C. 
Martínez-Rey, R. Marabini, J.M. Carazo. Fast and accurate conversion of atomic 
models into electron density maps. AIMS Biophysics, 2: 8-20 (2015)) and they are 
not identically distributed (it is well-known that Fourier components of low 
frequency have more energy than Fourier components of high frequency). Also, the 
assumption of iid $\xi_i$ variables is not well supported in EM in which we 
normally have noise before CTF (and consequently affected by CTF) and noise 
after CTF (which is not white, see C.O.S. Sorzano, S. Jonic, R. Núñez, N. 
Boisset, J.M. Carazo. Fast, robust and accurate determination of transmission 
electron microscopy contrast transfer function. Journal Structural
Biology 160: 249-262 (2007)). Despite of all these assumptions, which are not 
fulfilled in EM, it is better to have a method with incorrect assumptions but 
yielding useful results, than no method at all. And this seems to be the case 
with this algorithm. \\
\textbf{\textit{We agree with the reviewer, and to clarify this we have made a note 
about our assumptions in section 2.1. We also now cite the above
mentioned references. In this context, it makes sense to recall the famous saying of George Box: ``All models are wrong but some are useful." }}

\item In the article, it is not well remarked that the Epsilon covariance matrix 
is supposed to represent the covariance of the Fourier coefficients of 
projection images irrespective of their orientation. This assumption implies an 
increase of variance (due to the different orientations), in addition to the 
variance due to the different frequencies, which is not well explained in the 
article. \\
\textbf{\textit{We agree with the reviewer that this point was not well enough explained in the paper. 
We have added the required clarification about covariance due to different orientations to  Section 2.1,
please see the sentence beginning with "Since the clean images are 2D projections of the 3D molecule in different orientations...''}}

\item Eq. 13. Since there is an important source of noise affected by the CTF, a 
purely "whitening" filter would boost the noise close to the CTF zeroes, and it 
would not be possible to have $W \xi_i$ distributed as $\mathcal{N}(0,\sigma^2 I)$. How to 
construct the whitening filter should be better explained. \\
\textbf{\textit{The whitening filter is constructed by estimating the 2D isotropic 
power spectral density of noise. To do this, we first estimate using correlograms the 2D autocorrelation
of the corner pixels of the images which contain mostly noise and no signal. These corner
pixels are used to estimate the 1D autocorrelation, which is then extended to populate the 2D isotropic
autocorrelation. We have added a description of how $W$ is constructed in section 2.2.}}

\item It would be very interesting to see how the method deals in practice with 
contaminating images, further violating the method assumptions (for instance, 
images with just noise and no particles, particles with other particles in their 
surroundings, images with objects not being particles; this kind of images are 
easily obtained by automatic picking articles and they normally represent 
between $10-30\%$ of the collected data). \\
\textbf{\textit{Please see `Outlier Detection' in the `Improvements to the Manuscript' section.}}

\item Although maybe out of the scope of this paper, it would be very 
interesting to see if the restored particles can be used for 2D classification 
and even 3D reconstruction without a significant loss of structural information. 
At least, the authors could comment on this issues as future work. \\
\textbf{\textit{Indeed, this is a very important direction to pursue, and it 
is included in our plan for immediate future work.
We have added a few directions for future work in section 4.}}

\item In any case, the results shown by the authors are quite promising and the 
fact that EM images do not fulfill the theoretical assumptions of the method 
does not invalidate its results. The restored images are certainly closer to the 
ideal images than the acquired projections.\\
\textbf{\textit{Thanks, we completely agree.}}
\end{itemize}

\section*{Response to Reviewer 2}
\begin{itemize}
\item  People spend time looking at their particle images to discover outliers.  
It would be great if denoising helps this.  It would be useful in your 
simulations to show what outliers look like when denoised.  For example, what do 
pure-noise images look like when denoised?  Will they just look like $\mu$?  What 
if rare (a few $\%$ of total) images are just blobs, could one distinguish them?\\
\textbf{\textit{Please see `Outlier Detection' in the `Improvements to the Manuscript' section.}}

\item  Similarly, in the heterogeneous 70S dataset, do the denoised images 
demonstrate the heterogeneity?  That is, could one classify the denoised images 
correctly if one knew what the different underlying structures looked like?  
Apart from this issue, there seems little need to add this fourth dataset.\\
\textbf{\textit{The current CWF algorithm is not expected to be able to distinguish heterogeneous particles
based on their denoised images. We have added a comment to that effect in section 3.7. We show results with 2 datasets
each for new and old cameras simply to convince the readers of the applicability of CWF to experimental datasets.}}

\item  As ice thickness varies, the contrast in the particle images varies by as 
much as a factor of 2.  What would that do to the denoising?  I guess this is 
related to the question of what pure-noise images look like.\\
\textbf{\textit{The contrast factor due to variability in ice thickness can be modeled as an additional scalar parameter
and absorbed into the random variable $X$ in the simulation. CWF estimates the denoised image with its contrast included. To demonstrate this, 
we include varying contrast in the outlier detection experiment in section 3.7.}}

\item In a sense you are doing something similar to class averaging.  
Class-averaging is nice because it allows the user to discover classes of "bad" 
particles and delete all the underlying images from the dataset before 
reconstruction.  I'm trying to imagine how the use of the denoising procedure 
would fit into the workflow for SPR.  For example, would there be a way to look 
at the "neighborhood" of bad particles, say in the space of eigenvectors of 
$\Sigma$, to be able to exclude them?  Also, have you through this work discovered 
a better way to do class averaging in the first place?\\
\textbf{\textit{Indeed, this is a very important direction to pursue, and it 
is included in our plan for immediate future work.
Including CWF in the class averaging procedure is certainly at the top of our list for 
the next steps, as we remark in section 4.}}

\item  I'm left a little nervous by the eigenvalue thresholding procedures, not 
knowing what they are doing to the results.  Maybe you could give an example of 
the effect on the results of taking different sample sizes n, to show that the 
shrinkage procedure isn't introducing artifacts.\\
\textbf{\textit{Shrinkage was first shown
to improve covariance estimation by Stein \cite{stein1, stein2}, and is especially crucial in the high dimensional
setting \cite{donoho} such as in this paper.
We demonstrate the usefulness of the eigenvalue thresholding procedure in Figure 3, by showing
the improvement in accuracy of the covariance estimation when eigenvalue shrinkage is included. }}

\item  Maybe you should discuss briefly that you are not considering 
translations, just in-plane rotations in modeling the images.\\
\textbf{\textit{The algorithm does not assume centered images; however non-centered images
would lead to further variability and therefore an additional blurring effect. We have added a comment to that
effect at the beginning of section 3. }}

\item  Page 9, middle.  I'm confused by the statement "$p_k$ decreases with 
increasing k."   I'd expect the opposite.\\
\textbf{\textit{ $k$ is the angular frequency, and $p_k$ is the maximum radial frequency for a given $k$. 
Because the images are bandlimited, in order to avoid aliasing, for large angular frequencies we cannot allow large radial frequencies.
The exact relation is given in eqn. 13 of \cite{ffbspca}.}}

\item  Page 12, middle.  Please explain more what you mean by MSE.  Is it the 
norm of the raw image minus the CTF-filtered denoised version?\\
\textbf{\textit{Indeed, the MSE is the norm of the difference between the true, clean image and its denoised version. We have added a sentence
explicitly defining the MSE in the Results section.}}
\end{itemize}

\bibliography{resp_ref}

\end{document}

### 2017-02-08 elsevier_template/elsarticle-template.tex
JSB submitted version

\documentclass[review]{elsarticle}

\usepackage{lineno, hyperref}
\modulolinenumbers[5]

\journal{Journal of \LaTeX\ Templates}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author{Elsevier\fnref{myfootnote}}
\address{Radarweg 29, Amsterdam}
\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
\ead[url]{www.elsevier.com}

\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{support@elsevier.com}

\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
This template helps you to create a properly formatted \LaTeX\ manuscript.
\end{abstract}

\begin{keyword}
\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
\MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\linenumbers

\section{The Elsevier article class}

\paragraph{Installation} If the document class \emph{elsarticle} is not available on your computer, you can download and install the system package \emph{texlive-publishers} (Linux) or install the \LaTeX\ package \emph{elsarticle} using the package manager of your \TeX\ installation, which is typically \TeX\ Live or Mik\TeX.

\paragraph{Usage} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript. Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal. It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.

\paragraph{Functionality} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class. In addition, it features commands and options to format the
\begin{itemize}
\item document style
\item baselineskip
\item front matter
\item keywords and MSC codes
\item theorems, definitions and proofs
\item lables of enumerations
\item citation style and labeling.
\end{itemize}

\section{Front matter}

The author names and affiliations could be formatted in two ways:
\begin{enumerate}[(1)]
\item Group the authors per affiliation.
\item Use footnotes to indicate the affiliations.
\end{enumerate}
See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.

\section{Bibliography styles}

There are various bibliography styles available. You can select the style of your choice in the preamble of this document. These styles are Elsevier styles based on standard styles like Harvard and Vancouver. Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.

Here are two sample references: \cite{Feynman1963118,Dirac1953888}.

\section*{References}

\bibliography{mybibfile}

\end{document}

### 2017-02-08 siam_talk/nyu_bhamre.tex
JSB submitted version

\documentclass{beamer}
\usetheme{CambridgeUS}
\usefonttheme{serif} 
\usepackage[utf8]{inputenc}
\usepackage{default}
\usepackage{booktabs}
\usepackage{pifont}
\usepackage{lineno}
%\usepackage{beamerthemesplit}
%\usepackage{stmaryrd}
\usepackage{comment}
\usepackage{subfig}
\usepackage[percent]{overpic}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{movie15}
\usepackage{mathrsfs}
\usepackage{fixltx2e}
\newcommand{\argmin}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{min}}\;}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\begin{document}

\title{Denoising and Covariance Estimation of Cryo-EM Images} 
\institute{Princeton University\\
\it{Joint work with Teng Zhang (UCF) and Amit Singer (Princeton)}} 
\date{October 28, 2016} 
\author{Tejal Bhamre\\
} 
\begin{frame}
\titlepage 
\end{frame}

% 
% \begin{frame}{Overview}
% \tableofcontents
% \end{frame}
%

\section{Results on Experimental Datasets}

\begin{frame}<beamer>
\frametitle{Experimental data - 80S ribosome}

\begin{figure}[h]
\centering
{\begin{overpic}[width=0.5\textwidth]{jsb_fig_80s.eps}%
\put(8,77){\tiny Raw}
\put(22,77){\tiny Closest projection}
\put(59,77){\tiny TWF}
\put(83,77){\tiny CWF}
\end{overpic}
\label{}}
\label{fig:real80s}
\end{figure}
%  FEI
\begin{itemize}
 \item FALCON II 4k$\times$4k direct electron detector\\
 \item 105247 motion corrected, picked particle images of 360$\times$360 pixels
\end{itemize}
\end{frame}

\begin{frame}<beamer>
\frametitle{Experimental data - TRPV1}

\begin{figure}[h]
\centering
{\begin{overpic}[width=0.5\textwidth]{jsb_fig_tv.eps}%
\put(10,77){\tiny Raw}
\put(22,77){\tiny Closest projection}
\put(59,77){\tiny TWF}
\put(82,77){\tiny CWF}
\end{overpic}
\label{}}

\label{fig:trpv1}
\end{figure}
% 
\begin{itemize}
 \item K2 direct electron detector\\
 \item 35645 motion corrected, picked particle images of 256$\times$256 pixels
\end{itemize}
\end{frame}

\begin{frame}<beamer>
\frametitle{Experimental data -IP\textsubscript{3}R1}
\begin{figure}[h]
\centering
{\begin{overpic}[width=0.5\textwidth]{jsb_fig_ip3.eps}%
\put(10,77){\tiny Raw}
\put(22,77){\tiny Closest projection}
\put(57,77){\tiny TWF}
\put(83,77){\tiny CWF}
\end{overpic}
\label{}}
\label{fig:ip3}
\begin{itemize}
 \item Gatan 4k$\times$4k CCD\\
 \item 37382 picked particle images of 256$\times$256 pixels
\end{itemize}
\end{figure}
\end{frame}

\begin{frame}<beamer>
\frametitle{Experimental data - 70S ribosome}

\begin{figure}[h]
\centering
{\begin{overpic}[width=0.5\textwidth]{jsb_fig_70s.eps}%
\put(10,77){\tiny Raw}
\put(22,77){\tiny Closest projection}
\put(58,77){\tiny TWF}
\put(82,77){\tiny CWF}
\end{overpic}
\label{}}
\label{fig:real70s}
\end{figure}
\begin{itemize}
 \item TVIPS TEMCAM-F415 (4k x 4k) CCD\\
 \item 216517 picked particle images of 250$\times$250 pixels
\end{itemize}
\end{frame}

\section{Introduction}
\begin{frame}<beamer>
\setbeamercovered{transparent}
\frametitle{Cryo-EM Pipeline}
\begin{itemize}[<+->]
 \item Particle picking from micrographs
 \item \textbf{Preprocessing}
 \item \textbf{2D classification (Class averaging): inspect underlying particles, estimate viewing angles}
 \item 3D classification
 \item 3D refinement
\end{itemize}
\end{frame}

\begin{frame}<beamer>
\setbeamercovered{transparent}
\frametitle{Motivation}
\begin{itemize}[<+->]
 \item Visualization of underlying particles \alert{without class averaging}
 \item Image restoration (CTF correction and denoising) in a \alert{single step}
  \item \alert{Outlier detection}
\end{itemize}
\end{frame}

\begin{frame}<beamer>
\setbeamercovered{transparent}
\frametitle{CTF Correction}
\begin{figure}[]
%\centering
 \includegraphics[scale=0.4]{ctf2-ice.png}
 \footnote{\url{http://www.bio.brandeis.edu/~shaikh/lab/ctf.htm}}
\end{figure}

\begin{itemize}[<+->]
\item CTF suppresses information and inverts contrast
\item \alert{Challenge}: CTF cannot be trivially inverted (zero crossings)
\item Information lost from one defocus group could be recovered from another group that has different zero crossings.
\end{itemize}
\end{frame}

\begin{frame}<beamer>
\setbeamercovered{transparent}
\frametitle{Current Image Restoration Techniques}
\begin{itemize}[<+->]
 \item \textbf{Phase flipping + steerable PCA (sPCA)}: 
  \begin{itemize}
  \item  Flip sign of the Fourier
coefficients at frequencies for which the CTF is negative
   \item Preserves noise statistics
   \item Data adaptive basis: eigenvectors of the sample covariance matrix
   \item Phase flipping corrects only phases
  \end{itemize}
 \item \textbf{Traditional Wiener Filtering (TWF)}:
 \begin{itemize}
  \item  Corrects both phases and amplitudes
  \item Requires prior estimation of the spectral signal to
 noise ratio (SSNR)
  \item  Cannot restore information at zero crossings of the CTF
  \item Not in a data adaptive basis (restricted to Fourier basis)
 \end{itemize}
\end{itemize}
\end{frame}

\section{Covariance Wiener Filtering (CWF)}
\begin{frame}<beamer>
\setbeamercovered{transparent}
\frametitle{Covariance Wiener Filtering (CWF)}
\begin{itemize}[<+->]
 \item Estimate the CTF-corrected covariance matrix of the
underlying clean 2D projection images
\item Wiener filtering to solve the image restoration deconvolution problem
\item No averaging, act on each image separately
\item CTF correction and denoising in a single step
\end{itemize}
\end{frame}

\begin{frame}<beamer>
\frametitle{Covariance Wiener Filtering (CWF)}
\begin{table}[h!]
\small
  \centering
  \caption{Comparison of CTF Correction/Denoising Methods}
  \label{tab:table1}
  \begin{tabular}{cccc}
    \toprule
    Property & Phaseflip + sPCA & TWF & \alert{CWF}\\
    \midrule
    Applicable at preliminary stage & \cmark  & \cmark  & \cmark \\
    Data dependent basis & \cmark  &  \xmark & \cmark \\
    Correct both phases and amplitudes & \xmark  &  \cmark & \cmark \\
    CTF corrected covariance estimate & \textcolor{black}{\ding{51}}{\small\textcolor{black}{\kern-0.7em\ding{55}}}  &  \xmark & \cmark \\
 \bottomrule
  \end{tabular}
\end{table}
\end{frame}
% 
% \begin{frame}<beamer>
% \setbeamercovered{transparent}
% \frametitle{Fourier-Bessel Steerable Basis }
% \begin{itemize}[<+->]
% \item The population covariance matrix $\Sigma$ must be invariant under in-plane 
% rotation of the projection images
% \item Block diagonal in any steerable basis in which the 
% basis elements are outer products of radial functions and angular Fourier modes \footnote{\tiny{Fast Steerable Principal Component Analysis,
% Z. Zhao and Y. Shkolnisky and A. Singer, IEEE Transactions on Computational Imaging}}
% \item  CTF and the whitening filter
% are also block diagonal in the Fourier Bessel basis (radial isotropy)
% \item Suffices to 
% estimate each diagonal block of $\Sigma$, corresponding to the angular frequency $k$, separately
% \item \alert{Nearly unitary transformation}
% \end{itemize}
% \end{frame}

\begin{frame}<beamer>
\frametitle{Computational Complexity: Timings}
$n$ images of size $L \times L$\\
UNIX environment with 60 cores,
running at 2.3 GHz, with total RAM of 1.5TB

\begin{table}[t]
  \centering
  \caption{Timing in seconds}
  \begin{tabular}{ccccc}
    \toprule
    Dataset & $L$ & $n$ & Basis coeffs  & CWF \\
    \midrule
    TRPV1 & 256 & 35645 & 312s & 574s \\
    80s & 360 & 30000 &  731s & 385s \\
    IP3R1  & 256 & 37382 & 429s & 589s\\
    70s & 250 &  99979 & 1174s & 113s \\
 \bottomrule
  \end{tabular}
\end{table}
\end{frame}

\section{Results}
\begin{frame}<beamer>
\frametitle{Relative error of estimated clean images}

\begin{figure}[]
\centering
\subfloat[]{\begin{overpic}[width=0.5\linewidth]{mse_snr_6454.eps}%
\end{overpic}%
\label{fig:mse_snr}}
\subfloat[]{\begin{overpic}[width=0.5\linewidth]{mse_nims_6454.eps}
\end{overpic}
\label{fig:mse_nims}}
\caption{(a) {Fixed number of images}
(b) {Fixed SNR}
}
\end{figure}
\end{frame}

\begin{frame}<beamer>
\frametitle{Relative error of estimated covariance}
\begin{figure}
\centering
\includegraphics[width=0.6\linewidth]{cwf_shrinkage_compare.eps}
\caption{The estimator $\hat \Sigma$ can be shown to be consistent in the large sample limit
$n \to \infty$}
\label{fig:shrinkage}
\end{figure}
\end{frame}

\begin{frame}<beamer>
\frametitle{Simulations with white noise:  80S ribosome (EMDB-6454)}

\begin{figure}[]
\centering
\subfloat[SNR=$1$]{\begin{overpic}[width=0.4\linewidth]{compare_ims_snr1by1_6454.eps}
\put(6,52){\tiny Clean}
\put(32,52){\tiny Noisy}
\put(55,52){\tiny TWF}
\put(80,52){\tiny CWF}
\put(-32,35){\tiny Defocus=$1\mu m$}
\put(-32,10){\tiny Defocus=$4\mu m$}
\end{overpic}%
}
\vspace{-1mm}
\subfloat[SNR=$1/20$]{\begin{overpic}[width=0.4\linewidth]{compare_ims_snr1by20_6454.eps}%
%\put(5,40){Clean}
\end{overpic}%
\label{}}
\vspace{-1mm}
\subfloat[SNR=$1/40$]{\begin{overpic}[width=0.4\linewidth]{compare_ims_snr1by40_6454.eps}%
%\put(5,40){Clean}
\end{overpic}%
\label{}}
\vspace{-3mm}
\subfloat[SNR=$1/60$]{\begin{overpic}[width=0.4\linewidth]{compare_ims_snr1by60_6454.eps}%
%\put(5,40){Clean}
\end{overpic}%
\label{}}

\label{fig:ims_6454}
\end{figure}

\end{frame}

\begin{frame}<beamer>
\frametitle{Simulations with colored noise: 80S ribosome (EMDB-6454)}

\begin{figure}[]
\centering
\subfloat[SNR=$1$]{\begin{overpic}[width=0.4\linewidth]{compare_ims_col_snr1by1_6454.eps}
\put(6,52){\tiny Clean}
\put(32,52){\tiny Noisy}
\put(55,52){\tiny TWF}
\put(80,52){\tiny CWF}
\put(-32,35){\tiny Defocus=$1\mu m$}
\put(-32,10){\tiny Defocus=$4\mu m$}
\end{overpic}%
}
\vspace{-1mm}
\subfloat[SNR=$1/10$]{\begin{overpic}[width=0.4\linewidth]{compare_ims_col_snr1by10_6454.eps}%
\end{overpic}%
\label{}}
\vspace{-1mm}
\subfloat[SNR=$1/20$]{\begin{overpic}[width=0.4\linewidth]{compare_ims_col_snr1by20_6454.eps}%
\end{overpic}%
\label{}}

\label{fig:ims_6454_colored}
\end{figure}

\end{frame}

\begin{frame}<beamer>
\frametitle{Outlier Detection}
\setbeamercovered{transparent}
\begin{itemize}[<+->]
 \item  Significant amount of time is spent on discarding outliers by visual inspection
after the particle picking 
\item CWF:  automatic way to classify picked particles
\item Specimen particles at various depths in the ice layer: acquired
projection images can have different contrasts
\item  Outlier images typically have low contrast
after denoising: linear classifier
after CWF 
\end{itemize}
\end{frame}

\begin{frame}<beamer>
\frametitle{Outlier Detection: 80S ribosome (EMDB-6454)}
SNR=1/20 
$\alpha \in [0.75,1.5]$
$10\%$ images are pure noise
\begin{figure}[]
\centering
\subfloat[]{\begin{overpic}[width=0.34\linewidth]{raw_title.eps}%
\end{overpic}%
\label{fig:raw_outlier}}
\quad
\subfloat[]{\begin{overpic}[width=0.34\linewidth]{den_title.eps}
\end{overpic}
\label{fig:den_outlier}} \\
\subfloat[]{\begin{overpic}[width=0.1\linewidth]{mean_image.eps}%
\end{overpic}%
\label{fig:mean_image}}
\quad
\subfloat[]{\begin{overpic}[width=0.45\linewidth]{top_eigim.eps}
\end{overpic}
\label{fig:eigenims}}
\end{figure}
\end{frame}

\section{Work in Progress}
\begin{frame}<beamer>
\setbeamercovered{transparent}
\frametitle{Applications}
\begin{itemize}[<+->]
 \item Better class averages
 \item Covariance matrix: Orthogonal Replacement using Kam's method
 \item 3D reconstruction from denoised images, without class averaging
\end{itemize}
\end{frame}

\begin{frame}<beamer>
\setbeamercovered{transparent}
\frametitle{Resources}
\begin{itemize}
 \item Code: \url{https://github.com/PrincetonUniversity/cwf_denoise}
 \item Paper: Journal of Structural Biology: \url{10.1016/j.jsb.2016.04.013}
\end{itemize}
\end{frame}

\section{Methods}
\begin{frame}<beamer>
\frametitle{Pipeline}
\begin{itemize}
 \item Preprocessing
 \item Mean estimation
 \item Covariance Estimation
 \item Deconvolution

\end{itemize}

\end{frame}

\begin{frame}<beamer>
\frametitle{The Model: Real space}
Linear, weak phase approximation\\

\begin{equation}
 y_i = a_i \ast x_i + \epsilon_i, \quad i=1,2,\ldots,n
\label{eqn:model}
\end{equation}

$n$: number of images\\
$\ast$: convolution operation\\
$y_i$: noisy, CTF filtered $i$'th image in real space\\
$x_i$: underlying clean projection image in real space\\
$a_{i}$: the point spread function of the microscope\\
$\epsilon_i$: additive Gaussian noise that corrupts the image
\end{frame}

\begin{frame}<beamer>
\frametitle{The Model: Fourier space}

\begin{equation}
 Y_i = A_i X_i + \xi_i, \quad i=1,2,\ldots,n
\label{eqn:model_f}
\end{equation}

$A_i$: diagonal operator, whose
diagonal consists of the Fourier transform of the point spread function\\

$X_1,\dots,X_n$: vectors in $\mathbb{C}^p$, where $p$ is the number of pixels\\
i.i.d. samples from a distribution with mean $\mathbb{E}[\textbf{X}]=\mu$
and covariance $\mathbb{E}[(\textbf{X}-\mu)(\textbf{X}-\mu)^T]=\Sigma$\\

\textbf{\alert{``All models are wrong but some are useful" - George Box}}
\end{frame}

\begin{frame}<beamer>
\frametitle{The Model}
\begin{equation}
\mathbb{E}[\textbf{Y}_i]=A_i \mathbb{E}[\textbf{X}_i], \quad i=1,2,\ldots,n.
\label{eqn:exp_y}
\end{equation}

\begin{equation}
\begin{aligned}
\mathbb{E}[(\textbf{Y}_i-\mathbb{E}[\textbf{Y}_i])(\textbf{Y}_i-\mathbb{E}[\textbf{Y}_i])^T] 
&= \mathbb{E} [A_i(\textbf{X}_i-\mu)(\textbf{X}_i-\mu)^T A_i^T] + \sigma^2I \\
&=  A_i \Sigma A_i^T + \sigma^2I .
\end{aligned}
\label{eqn:expectation_eq}
\end{equation}

Relates the second order statistics of the noisy images with the 
population covariance $\Sigma$ of the clean images
\end{frame}

%%%%%

\begin{frame}<beamer>
\frametitle{Mean Estimation}

\begin{equation}
 \hat\mu = \argmin{\mu} \sum_{i=1}^n||(Y_i-A_i\mu)||_2^2 + \lambda||\mu||_2^2
\end{equation}

% % 
\begin{equation}
 \hat\mu = (\sum_{i=1}^n A_i^T A_i + \lambda I)^{-1}(\sum_{i=1}^n 
A_i^T Y_i).
\label{eq:ls_mean_sol}
\end{equation}
\end{frame}

\begin{frame}<beamer>
\frametitle{Covariance Estimation}
\begin{equation}
\begin{aligned}
\hat\Sigma 
&= \argmin{\Sigma} \sum_{i=1}^n || (Y_i - \mathbb{E}[\textbf{Y}_i]) (Y_i - \mathbb{E}[\textbf{Y}_i])^T
- (A_i \Sigma A_i^T + \sigma^2 I)||_F^2 \\
&= \argmin{\Sigma} \sum_{i=1}^n || A_i\Sigma A_i^T + \sigma^2 I - C_i  ||_F^2 
\end{aligned}
\label{eqn:ls1}
\end{equation}
where $C_i=(Y_i - A_i \mu) (Y_i - A_i \mu)^T$ and $||.||_F$ is the Frobenius matrix norm. 
\end{frame}

\begin{frame}<beamer>
\frametitle{Solving using Conjugate Gradient}
System of linear equations for the elements of the matrix $\hat \Sigma$

\begin{equation}
\begin{aligned}
\sum_{i=1}^n  A_i^T  A_i \hat \Sigma A_i^T A_i
&= \sum_{i=1}^n A_i^T C_i A_i - \sum_{i=1}^n \sigma^2 A_i^T A_i 
\end{aligned}
\label{eqn:normal_white}
\end{equation}
\frametitle{Solving using Conjugate Gradient}
\begin{equation}
\begin{aligned}
L(\hat\Sigma) 
&=  B 
\label{eqn:cg}
\end{aligned}
\end{equation}
where $L:\mathbb{R}^{p\times p} \to \mathbb{R}^{p\times p}$ is the linear operator acting on $\hat{\Sigma}$ defined by the left hand side of eqn. \ref{eqn:normal_white}, and $B$ is the right hand side.

\begin{itemize}
 \item Direct inversion of this linear system is slow for large image sizes
 \item Applying $L$ only involves matrix multiplications: fast!
 \item Conjugate gradient 
\end{itemize}

\end{frame}

\begin{frame}<beamer>
\frametitle{Wiener Filtering}

\begin{itemize}
\item White noise: estimate $X_i$ as
\begin{equation}
\hat X_i = (I-H_iA_{i})\hat\mu + H_iY_i 
\end{equation}
where $H_i = \hat \Sigma A_{i}^T ( A_{i} \hat \Sigma A_{i}^T + \sigma^2 
I)^{-1} $ is the linear Wiener filter  
\item Colored noise: estimate $X_i$ as
\begin{equation}
\hat X_i = (I-H_iWA_{i})\hat\mu + H_iY_i 
\end{equation}
with $H_i = \hat \Sigma A_{i}^T W^T (W A_{i} \hat \Sigma A_{i}^T W^T 
+ \sigma^2 I)^{-1}$
\end{itemize}
\end{frame}

\begin{frame}<beamer>
\setbeamercovered{transparent}
\frametitle{Acknowledgement}
\begin{itemize}
 \item  NIGMS, AFOSR, Simons
Foundation, Moore Foundation Data-Driven Discovery Investigator Award
\item  Datasets: Maofu Liao, Sjors Scheres, Irina Serysheva and Joachim Frank
\item Yoel Shkolniskly, Fred Sigworth, Zhizhen Zhao, Joakim And\'en
\end{itemize}
\end{frame}
\end{document}

### 2017-02-08 siam_talk/siam_bhamre.tex
JSB submitted version

--- 2017-02-08 siam_talk/nyu_bhamre.tex
@@
 \it{Joint work with Teng Zhang (UCF) and Amit Singer (Princeton)}} 
-\date{October 28, 2016} 
+\date{June 10, 2016} 
 \author{Tejal Bhamre\\
@@
 \setbeamercovered{transparent}
+\frametitle{CTF Correction}
+
+\begin{itemize}
+\item Do nothing (Denial) 
+\item Correct Fourier phases but not amplitudes
+\item Correct both Fourier phases and amplitudes 
+\end{itemize}
+\end{frame}
+
+\begin{frame}<beamer>
+\setbeamercovered{transparent}
 \frametitle{Current Image Restoration Techniques}
@@
 \end{frame}
-% 
-% \begin{frame}<beamer>
-% \setbeamercovered{transparent}
-% \frametitle{Fourier-Bessel Steerable Basis }
-% \begin{itemize}[<+->]
-% \item The population covariance matrix $\Sigma$ must be invariant under in-plane 
-% rotation of the projection images
-% \item Block diagonal in any steerable basis in which the 
-% basis elements are outer products of radial functions and angular Fourier modes \footnote{\tiny{Fast Steerable Principal Component Analysis,
-% Z. Zhao and Y. Shkolnisky and A. Singer, IEEE Transactions on Computational Imaging}}
-% \item  CTF and the whitening filter
-% are also block diagonal in the Fourier Bessel basis (radial isotropy)
-% \item Suffices to 
-% estimate each diagonal block of $\Sigma$, corresponding to the angular frequency $k$, separately
-% \item \alert{Nearly unitary transformation}
-% \end{itemize}
-% \end{frame}
+
+\begin{frame}<beamer>
+\setbeamercovered{transparent}
+\frametitle{Fourier-Bessel Steerable Basis }
+\begin{itemize}[<+->]
+\item The population covariance matrix $\Sigma$ must be invariant under in-plane 
+rotation of the projection images
+\item Block diagonal in any steerable basis in which the 
+basis elements are outer products of radial functions and angular Fourier modes \footnote{\tiny{Fast Steerable Principal Component Analysis,
+Z. Zhao and Y. Shkolnisky and A. Singer, IEEE Transactions on Computational Imaging}}
+\item  CTF and the whitening filter
+are also block diagonal in the Fourier Bessel basis (radial isotropy)
+\item Suffices to 
+estimate each diagonal block of $\Sigma$, corresponding to the angular frequency $k$, separately
+\item \alert{Nearly unitary transformation}
+\end{itemize}
+\end{frame}
+
+\section{Methods}
+\begin{frame}<beamer>
+\frametitle{The Model: Real space}
+Linear, weak phase approximation\\
+
+\begin{equation}
+ y_i = a_i \ast x_i + \epsilon_i, \quad i=1,2,\ldots,n
+\label{eqn:model}
+\end{equation}
+
+$n$: number of images\\
+$\ast$: convolution operation\\
+$y_i$: noisy, CTF filtered $i$'th image in real space\\
+$x_i$: underlying clean projection image in real space\\
+$a_{i}$: the point spread function of the microscope\\
+$\epsilon_i$: additive Gaussian noise that corrupts the image
+\end{frame}
+
+\begin{frame}<beamer>
+\frametitle{The Model: Fourier space}
+
+\begin{equation}
+ Y_i = A_i X_i + \xi_i, \quad i=1,2,\ldots,n
+\label{eqn:model_f}
+\end{equation}
+
+$A_i$: diagonal operator, whose
+diagonal consists of the Fourier transform of the point spread function\\
+
+$X_1,\dots,X_n$: vectors in $\mathbb{C}^p$, where $p$ is the number of pixels\\
+i.i.d. samples from a distribution with mean $\mathbb{E}[\textbf{X}]=\mu$
+and covariance $\mathbb{E}[(\textbf{X}-\mu)(\textbf{X}-\mu)^T]=\Sigma$\\
+
+\textbf{\alert{``All models are wrong but some are useful" - George Box}}
+\end{frame}
+
+\begin{frame}<beamer>
+\frametitle{The Model}
+\begin{equation}
+\mathbb{E}[\textbf{Y}_i]=A_i \mathbb{E}[\textbf{X}_i], \quad i=1,2,\ldots,n.
+\label{eqn:exp_y}
+\end{equation}
+
+\begin{equation}
+\begin{aligned}
+\mathbb{E}[(\textbf{Y}_i-\mathbb{E}[\textbf{Y}_i])(\textbf{Y}_i-\mathbb{E}[\textbf{Y}_i])^T] 
+&= \mathbb{E} [A_i(\textbf{X}_i-\mu)(\textbf{X}_i-\mu)^T A_i^T] + \sigma^2I \\
+&=  A_i \Sigma A_i^T + \sigma^2I .
+\end{aligned}
+\label{eqn:expectation_eq}
+\end{equation}
+
+Relates the second order statistics of the noisy images with the 
+population covariance $\Sigma$ of the clean images
+\end{frame}
+
+%%%%%
+
+\begin{frame}<beamer>
+\frametitle{Mean Estimation}
+
+\begin{equation}
+ \hat\mu = \argmin{\mu} \sum_{i=1}^n||(Y_i-A_i\mu)||_2^2 + \lambda||\mu||_2^2
+\end{equation}
+
+% % 
+\begin{equation}
+ \hat\mu = (\sum_{i=1}^n A_i^T A_i + \lambda I)^{-1}(\sum_{i=1}^n 
+A_i^T Y_i).
+\label{eq:ls_mean_sol}
+\end{equation}
+\end{frame}
+
+\begin{frame}<beamer>
+\frametitle{Covariance Estimation}
+\begin{equation}
+\begin{aligned}
+\hat\Sigma 
+&= \argmin{\Sigma} \sum_{i=1}^n || (Y_i - \mathbb{E}[\textbf{Y}_i]) (Y_i - \mathbb{E}[\textbf{Y}_i])^T
+- (A_i \Sigma A_i^T + \sigma^2 I)||_F^2 \\
+&= \argmin{\Sigma} \sum_{i=1}^n || A_i\Sigma A_i^T + \sigma^2 I - C_i  ||_F^2 
+\end{aligned}
+\label{eqn:ls1}
+\end{equation}
+where $C_i=(Y_i - A_i \mu) (Y_i - A_i \mu)^T$ and $||.||_F$ is the Frobenius matrix norm. 
+\end{frame}
+
+\begin{frame}<beamer>
+\frametitle{Solving using Conjugate Gradient}
+System of linear equations for the elements of the matrix $\hat \Sigma$
+
+\begin{equation}
+\begin{aligned}
+\sum_{i=1}^n  A_i^T  A_i \hat \Sigma A_i^T A_i
+&= \sum_{i=1}^n A_i^T C_i A_i - \sum_{i=1}^n \sigma^2 A_i^T A_i 
+\end{aligned}
+\label{eqn:normal_white}
+\end{equation}
+\frametitle{Solving using Conjugate Gradient}
+\begin{equation}
+\begin{aligned}
+L(\hat\Sigma) 
+&=  B 
+\label{eqn:cg}
+\end{aligned}
+\end{equation}
+where $L:\mathbb{R}^{p\times p} \to \mathbb{R}^{p\times p}$ is the linear operator acting on $\hat{\Sigma}$ defined by the left hand side of eqn.
+\ref{eqn:normal_white}, and $B$ is the right hand side.
+
+\begin{itemize}
+ \item Direct inversion of this linear system is slow for large image sizes
+ \item Applying $L$ only involves matrix multiplications: fast!
+ \item Conjugate gradient 
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}<beamer>
+\frametitle{Eigenvalue Thresholding}
+\setbeamercovered{transparent}
+\begin{itemize}[<+->]
+ \item  $L(\hat{\Sigma})$ is a PSD matrix whenever $\hat{\Sigma}$ is PSD (as 
+a sum of PSD matrices)
+\item  $B$ may not necessarily be PSD due to finite 
+sample fluctuations ( $n$ is finite)
+\item  Project 
+$B$ onto the cone of PSD matrices
+\item  Compute the spectral 
+decomposition of $B$ and set all negative eigenvalues to 0  (eigenvalue thresholding)
+\item $n \gg p$
+\end{itemize}
+\end{frame}
+
+\begin{frame}<beamer>
+\frametitle{Eigenvalue Shrinkage: Spiked Covariance Model}
+\setbeamercovered{transparent}
+Analyze $B$ when $X_i=0$ for all $i$ (input images 
+are white noise images)
+
+\begin{itemize}
+ \item  \begin{equation}
+M = \sum_{i=1}^n A_i^T C_i A_i = \sum_{i=1}^n A_i^T Y_i Y_i^T A_i.
+\end{equation}
+\item  $\mathbb{E}[M] = \sigma^2 \sum_{i=1}^n A_i^T A_i$
+and $B = M - \mathbb{E}[M]$
+\item  $S = (\mathbb{E}[M])^{1/2}$, i.e. 
+$S$ is PSD and $\mathbb{E}[M]=S^2$
+\item  \begin{equation}
+ S^{-1} L(\hat\Sigma)  S^{-1} = S^{-1}(M - \mathbb{E}[M]) S^{-1} = S^{-1} M S^{-1} - I .
+\label{eqn:pop1}
+\end{equation}
+\end{itemize}
+\end{frame}
+
+\begin{frame}<beamer>
+\frametitle{Eigenvalue Shrinkage: Spiked Covariance Model}
+\setbeamercovered{transparent}
+\begin{itemize}[<+->]
+\item $S^{-1}MS^{-1}$ can be viewed as a sample covariance matrix of
+$n$ vectors in $ \mathbb{R}^p$ whose population covariance is the identity matrix
+\item Eigenvalues corresponding to the signal 
+can only be detected if they reside outside of the support of the  Mar\v{c}enko Pastur (MP)
+distribution
+\item Kritchman Nadler (KN) rank estimation to determine the number of eigenvalues 
+corresponding to the signal \footnote{\tiny{Determining the number of components in a factor model from limited noisy data, Shira Kritchman and Boaz Nadler,
+Chemometrics and Intelligent Laboratory Systems}}
+\item Apply operator norm eigenvalue 
+shrinkage procedure (Donoho et al.) to those eigenvalues, while setting all other eigenvalues to 
+$0$ \footnote{\tiny{Optimal shrinkage of eigenvalues in the spiked covariance model,  David Donoho, Matan Gavish and I M Johnstone,
+\url{arxiv.org/abs/1311.0851}}}
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}<beamer>
+\frametitle{Wiener Filtering}
+
+\begin{itemize}
+\item White noise: estimate $X_i$ as
+\begin{equation}
+\hat X_i = (I-H_iA_{i})\hat\mu + H_iY_i 
+\end{equation}
+where $H_i = \hat \Sigma A_{i}^T ( A_{i} \hat \Sigma A_{i}^T + \sigma^2 
+I)^{-1} $ is the linear Wiener filter  
+\item Colored noise: estimate $X_i$ as
+\begin{equation}
+\hat X_i = (I-H_iWA_{i})\hat\mu + H_iY_i 
+\end{equation}
+with $H_i = \hat \Sigma A_{i}^T W^T (W A_{i} \hat \Sigma A_{i}^T W^T 
++ \sigma^2 I)^{-1}$
+\end{itemize}
+\end{frame}
+
+\begin{frame}<beamer>
+\frametitle{Computational Complexity}
+$O(TDL^4 + nL^3)$, where $T$ is the number of conjugate gradient iterations
+\begin{itemize}
+\item $D$ defocus groups with $d_i$ images in group $i$
+\item Images of size $L \times L$
+\item \textit{n} images
+\end{itemize}
+\end{frame}
 
@@
 projection images can have different contrasts
+\item 
+\begin{equation}
+ Y_i = \alpha_i A_i X_i + \xi_i, \quad i=1,2,\ldots,n
+\label{eqn:contrast}
+\end{equation}
+\item Absorb $\alpha$ into $\textbf{X}$ and estimate $\alpha_i X_i$ 
 \item  Outlier images typically have low contrast
@@
 \setbeamercovered{transparent}
-\frametitle{Applications}
+\frametitle{Current and Future Work}
 \begin{itemize}[<+->]
@@
 
-\section{Methods}
-\begin{frame}<beamer>
-\frametitle{Pipeline}
-\begin{itemize}
- \item Preprocessing
- \item Mean estimation
- \item Covariance Estimation
- \item Deconvolution
-
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}<beamer>
-\frametitle{The Model: Real space}
-Linear, weak phase approximation\\
-
-\begin{equation}
- y_i = a_i \ast x_i + \epsilon_i, \quad i=1,2,\ldots,n
-\label{eqn:model}
-\end{equation}
-
-$n$: number of images\\
-$\ast$: convolution operation\\
-$y_i$: noisy, CTF filtered $i$'th image in real space\\
-$x_i$: underlying clean projection image in real space\\
-$a_{i}$: the point spread function of the microscope\\
-$\epsilon_i$: additive Gaussian noise that corrupts the image
-\end{frame}
-
-\begin{frame}<beamer>
-\frametitle{The Model: Fourier space}
-
-\begin{equation}
- Y_i = A_i X_i + \xi_i, \quad i=1,2,\ldots,n
-\label{eqn:model_f}
-\end{equation}
-
-$A_i$: diagonal operator, whose
-diagonal consists of the Fourier transform of the point spread function\\
-
-$X_1,\dots,X_n$: vectors in $\mathbb{C}^p$, where $p$ is the number of pixels\\
-i.i.d. samples from a distribution with mean $\mathbb{E}[\textbf{X}]=\mu$
-and covariance $\mathbb{E}[(\textbf{X}-\mu)(\textbf{X}-\mu)^T]=\Sigma$\\
-
-\textbf{\alert{``All models are wrong but some are useful" - George Box}}
-\end{frame}
-
-\begin{frame}<beamer>
-\frametitle{The Model}
-\begin{equation}
-\mathbb{E}[\textbf{Y}_i]=A_i \mathbb{E}[\textbf{X}_i], \quad i=1,2,\ldots,n.
-\label{eqn:exp_y}
-\end{equation}
-
-\begin{equation}
-\begin{aligned}
-\mathbb{E}[(\textbf{Y}_i-\mathbb{E}[\textbf{Y}_i])(\textbf{Y}_i-\mathbb{E}[\textbf{Y}_i])^T] 
-&= \mathbb{E} [A_i(\textbf{X}_i-\mu)(\textbf{X}_i-\mu)^T A_i^T] + \sigma^2I \\
-&=  A_i \Sigma A_i^T + \sigma^2I .
-\end{aligned}
-\label{eqn:expectation_eq}
-\end{equation}
-
-Relates the second order statistics of the noisy images with the 
-population covariance $\Sigma$ of the clean images
-\end{frame}
-
-%%%%%
-
-\begin{frame}<beamer>
-\frametitle{Mean Estimation}
-
-\begin{equation}
- \hat\mu = \argmin{\mu} \sum_{i=1}^n||(Y_i-A_i\mu)||_2^2 + \lambda||\mu||_2^2
-\end{equation}
-
-% % 
-\begin{equation}
- \hat\mu = (\sum_{i=1}^n A_i^T A_i + \lambda I)^{-1}(\sum_{i=1}^n 
-A_i^T Y_i).
-\label{eq:ls_mean_sol}
-\end{equation}
-\end{frame}
-
-\begin{frame}<beamer>
-\frametitle{Covariance Estimation}
-\begin{equation}
-\begin{aligned}
-\hat\Sigma 
-&= \argmin{\Sigma} \sum_{i=1}^n || (Y_i - \mathbb{E}[\textbf{Y}_i]) (Y_i - \mathbb{E}[\textbf{Y}_i])^T
-- (A_i \Sigma A_i^T + \sigma^2 I)||_F^2 \\
-&= \argmin{\Sigma} \sum_{i=1}^n || A_i\Sigma A_i^T + \sigma^2 I - C_i  ||_F^2 
-\end{aligned}
-\label{eqn:ls1}
-\end{equation}
-where $C_i=(Y_i - A_i \mu) (Y_i - A_i \mu)^T$ and $||.||_F$ is the Frobenius matrix norm. 
-\end{frame}
-
-\begin{frame}<beamer>
-\frametitle{Solving using Conjugate Gradient}
-System of linear equations for the elements of the matrix $\hat \Sigma$
-
-\begin{equation}
-\begin{aligned}
-\sum_{i=1}^n  A_i^T  A_i \hat \Sigma A_i^T A_i
-&= \sum_{i=1}^n A_i^T C_i A_i - \sum_{i=1}^n \sigma^2 A_i^T A_i 
-\end{aligned}
-\label{eqn:normal_white}
-\end{equation}
-\frametitle{Solving using Conjugate Gradient}
-\begin{equation}
-\begin{aligned}
-L(\hat\Sigma) 
-&=  B 
-\label{eqn:cg}
-\end{aligned}
-\end{equation}
-where $L:\mathbb{R}^{p\times p} \to \mathbb{R}^{p\times p}$ is the linear operator acting on $\hat{\Sigma}$ defined by the left hand side of eqn.
-\ref{eqn:normal_white}, and $B$ is the right hand side.
-
-\begin{itemize}
- \item Direct inversion of this linear system is slow for large image sizes
- \item Applying $L$ only involves matrix multiplications: fast!
- \item Conjugate gradient 
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}<beamer>
-\frametitle{Wiener Filtering}
-
-\begin{itemize}
-\item White noise: estimate $X_i$ as
-\begin{equation}
-\hat X_i = (I-H_iA_{i})\hat\mu + H_iY_i 
-\end{equation}
-where $H_i = \hat \Sigma A_{i}^T ( A_{i} \hat \Sigma A_{i}^T + \sigma^2 
-I)^{-1} $ is the linear Wiener filter  
-\item Colored noise: estimate $X_i$ as
-\begin{equation}
-\hat X_i = (I-H_iWA_{i})\hat\mu + H_iY_i 
-\end{equation}
-with $H_i = \hat \Sigma A_{i}^T W^T (W A_{i} \hat \Sigma A_{i}^T W^T 
-+ \sigma^2 I)^{-1}$
-\end{itemize}
-\end{frame}
-
 \begin{frame}<beamer>
@@
 Foundation, Moore Foundation Data-Driven Discovery Investigator Award
-\item  Datasets: Maofu Liao, Sjors Scheres, Irina Serysheva and Joachim Frank
+\item  Maofu Liao, Sjors Scheres, Irina Serysheva and Joachim Frank
 \item Yoel Shkolniskly, Fred Sigworth, Zhizhen Zhao, Joakim And\'en

### 2017-02-10 echo_talk.tex
Echo talk

--- 2017-02-08 siam_talk/siam_bhamre.tex
@@
 %
+
+\section{What is cryo electron microscopy?}
+
+\begin{frame}<beamer>
+\frametitle{Why cryo-EM?}
+
+%  FEI
+\begin{itemize}
+ \item 3D structure of macromolecules - understand function - drug discovery for cancer, AIDS, HIV, etc.
+ \item X ray crystallography: many proteins, viruses resistant to crystallization (HIV virus)
+ \item Cryo-EM: native state
+ \item Can study structures impossible before!
+\end{itemize}
+\end{frame}
+
+\begin{frame}<beamer>
+\frametitle{Nature Method of the Year 2015}
+\includegraphics[width=0.5\textwidth]{nature.png}%
+\end{frame}
+
+\begin{frame}<beamer>
+\frametitle{Cancer}
+\includegraphics[width=0.5\textwidth]{cancer.png}%
+\end{frame}
+
+\begin{frame}<beamer>
+\frametitle{Zika}
+\includegraphics[width=0.5\textwidth]{zika.png}%
+\end{frame}
+
+\section{Cryo-EM Pipeline}
+\begin{frame}
+\frametitle{3D from 2D}
+%  FEI
+\begin{itemize}
+ \item Flash freeze sample in an ice layer
+ \item Image with an electron microscope
+ \item 2D images at unkown orientations -> 3D structure
+ \item Sounds like Structure from Motion (SfM)?
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Sounds like SfM?}
+%  FEI
+\includegraphics[width=0.5\textwidth]{zika.png}%
+\begin{itemize}
+  \item Can't detect features from raw images, unlike SfM
+ \item Need new algorithms to handle noise level!
+\end{itemize}
+\end{frame}
 

### 2017-02-10 echo_talk.tex
Removed timings slide from echo talk

@@
 \begin{frame}
+\frametitle{3D from 2D}
+%  FEI
+\begin{itemize}
+ \item Insert pipeline flowchart
+\end{itemize}
+\end{frame}
+
+\begin{frame}
 \frametitle{Sounds like SfM?}
@@
  \item Need new algorithms to handle noise level!
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Cryo-EM Pipeline}
+%  FEI
+\includegraphics[width=0.5\textwidth]{zika.png}%
+\begin{itemize}
+  \item Can't detect features from raw images, unlike SfM
+ \item Need new algorithms to handle noise level!
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Open Source Software}
+%  FEI
+\includegraphics[width=0.5\textwidth]{zika.png}%
+\begin{itemize}
+  \item Insert ASPIRE image
+  \item Open source software toolbox: ASPIRE
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Today's talk}
+%  FEI
+\includegraphics[width=0.5\textwidth]{zika.png}%
+\begin{itemize}
+  \item Pipeline figure: highlight denoising, class averaging
 \end{itemize}
@@
 \end{itemize}
-\end{frame}
-
-\begin{frame}<beamer>
-\frametitle{Computational Complexity: Timings}
-$n$ images of size $L \times L$\\
 UNIX environment with 60 cores,
@@
 
-\begin{table}[t]
-  \centering
-  \caption{Timing in seconds}
-  \begin{tabular}{ccccc}
-    \toprule
-    Dataset & $L$ & $n$ & Basis coeffs  & CWF \\
-    \midrule
-    TRPV1 & 256 & 35645 & 312s & 574s \\
-    80s & 360 & 30000 &  731s & 385s \\
-    IP3R1  & 256 & 37382 & 429s & 589s\\
-    70s & 250 &  99979 & 1174s & 113s \\
- \bottomrule
-  \end{tabular}
-\end{table}
 \end{frame}

### 2017-02-10 echo_talk.tex
Mah dist slide echo talk

@@
 
-\section{Work in Progress}
+\section{Applications and Extensions}
 \begin{frame}<beamer>
 \setbeamercovered{transparent}
-\frametitle{Current and Future Work}
+\frametitle{Applications}
 \begin{itemize}[<+->]
  \item Better class averages
- \item Covariance matrix: Orthogonal Replacement using Kam's method
  \item 3D reconstruction from denoised images, without class averaging
@@
 \setbeamercovered{transparent}
-\frametitle{Resources}
-\begin{itemize}
- \item Code: \url{https://github.com/PrincetonUniversity/cwf_denoise}
- \item Paper: Journal of Structural Biology: \url{10.1016/j.jsb.2016.04.013}
+\frametitle{Better 2D classification}
+\begin{itemize}[<+->]
+ \item Mahalanobis distance pipeline figure from paper
 \end{itemize}
@@
 
-\begin{frame}<beamer>
-\setbeamercovered{transparent}
-\frametitle{Acknowledgement}
-\begin{itemize}
- \item  NIGMS, AFOSR, Simons
-Foundation, Moore Foundation Data-Driven Discovery Investigator Award
-\item  Maofu Liao, Sjors Scheres, Irina Serysheva and Joachim Frank
-\item Yoel Shkolniskly, Fred Sigworth, Zhizhen Zhao, Joakim And\'en
-\end{itemize}
-\end{frame}
 \end{document}
